High-Fidelity Counterfactual Explanations through Conformal Prediction

### 2023-04-24 notebooks/mnist.qmd
mnist

```{julia}
# Train models:
function _train(model, X=X, y=labels; cov=.95, method=:simple_inductive)
    conf_model = conformal_model(jem_ens; method=method, coverage=cov)
    mach = machine(conf_model, X, y)
    fit!(mach)
    M = ECCCo.ConformalModel(mach.model, mach.fitresult)
    return M
end
model_dict = Dict(mod_name => _train(mod) for (mod_name, mod) in models)
```

```{julia}
if mach.model.model isa JointEnergyClassifier
    sampler = mach.model.model.jem.sampler
else
    K = length(counterfactual_data.y_levels)
    input_size = size(selectdim(counterfactual_data.X, ndims(counterfactual_data.X), 1))
    𝒟x = Uniform(extrema(counterfactual_data.X)...)
    𝒟y = Categorical(ones(K) ./ K)
    sampler = ConditionalSampler(𝒟x, 𝒟y; input_size=input_size)
end
opt = ImproperSGLD()
f(x) = logits(M, x)

n_iter = 200
_w = 1500
plts = []
neach = 10
for i in 1:10
    x = sampler(f, opt; niter=n_iter, n_samples=neach, y=i)
    plts_i = []
    for j in 1:size(x, 2)
        xj = x[:,j]
        xj = reshape(xj, (n_digits, n_digits))
        plts_i = [plts_i..., Plots.heatmap(rotl90(xj), axis=nothing, cb=false)]
    end
    plt = Plots.plot(plts_i..., size=(_w,0.10*_w), layout=(1,10))
    plts = [plts..., plt]
end
plt = Plots.plot(plts..., size=(_w,_w), layout=(10,1))
display(plt)

```

```{julia}
test_data = load_mnist_test()
test_data.X = pre_process.(test_data.X)
f1 = CounterfactualExplanations.Models.model_evaluation(M, test_data)
println("F1 score (test): $(round(f1,digits=3))")
```

>>>>

```{julia}
# Train models:
function _train(model, X=X, y=labels; cov=.95, method=:simple_inductive, mod_name="model")
    conf_model = conformal_model(model; method=method, coverage=cov)
    mach = machine(conf_model, X, y)
    @info "Begin training $mod_name."
    fit!(mach)
    @info "Finished training $mod_name."
    M = ECCCo.ConformalModel(mach.model, mach.fitresult)
    return M
end
model_dict = Dict(mod_name => _train(mod; mod_name=mod_name) for (mod_name, mod) in models)
Serialization.serialize(joinpath(output_path,"mnist_models.jls"), model_dict)
```

```{julia}
for (mod_name, mod) in model_dict

    # Plot:
    if mod.model.model isa JointEnergyClassifier
        sampler = mod.model.model.jem.sampler
    elseif mod.model.model.model isa JointEnergyClassifier
        sampler = mod.model.model.model.jem.sampler
    else
        K = length(counterfactual_data.y_levels)
        input_size = size(selectdim(counterfactual_data.X, ndims(counterfactual_data.X), 1))
        𝒟x = Uniform(extrema(counterfactual_data.X)...)
        𝒟y = Categorical(ones(K) ./ K)
        sampler = ConditionalSampler(𝒟x, 𝒟y; input_size=input_size)
    end
    opt = ImproperSGLD()
    f(x) = logits(mod, x)

    n_iter = 200
    _w = 1500
    plts = []
    neach = 10
    for i in 1:10
        x = sampler(f, opt; niter=n_iter, n_samples=neach, y=i)
        plts_i = []
        for j in 1:size(x, 2)
            xj = x[:,j]
            xj = reshape(xj, (n_digits, n_digits))
            plts_i = [plts_i..., Plots.heatmap(rotl90(xj), axis=nothing, cb=false)]
        end
        plt = Plots.plot(plts_i..., size=(_w,0.10*_w), layout=(1,10))
        plts = [plts..., plt]
    end
    plt = Plots.plot(plts..., size=(_w,_w), layout=(10,1), plot_title=mod_name)
    display(plt)

    # Test performance:
    test_data = load_mnist_test()
    test_data.X = pre_process.(test_data.X)
    f1 = CounterfactualExplanations.Models.model_evaluation(M, test_data)
    println("F1 score (test): $(round(f1,digits=3))")
end
```
